Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2017-07-21 paper.tex
edits throughout

The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation \cite{Harlan1992}.
However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
An important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.

@@

Over the last few years, several high-profile publications based on genome-wide data have documented introgression between crops and their wild relatives outside putative domestication centers.
Recent empirical studies have revealed that introgression has occurred in many of the world's most important crops (\textbf{Table 2}).

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter referred to as \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002single, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
%The following is not fully clear.  Cut or revise.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
Height differences were not detected under lowland conditions.
%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

@@

Populations of \emph{mexicana} cannot be found outside of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
%Several questions regarding \emph{mexicana} introgression yet remain.
%Has the introgression from \emph{mexicana} spread to multiple highland populations? 
%How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
%On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al. (2013) \citep{hufford2013}.
%The signal of introgression is absent for the other three populations.
%On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
%More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry. 
%The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects. 
%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
%The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
%Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
%Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.

>>>>

The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation \cite{Harlan1992}.
However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
An additional important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.

@@

Genome-wide data from extensive samples of crops and their wild relatives, in combination with the new methods described above, have recently allowed detailed analysis of wild-to-crop introgression in some of the world's most important crops (\textbf{Table 2}).
Below we present a summary of findings from maize, barley, and rice, three promising systems in which introgression from wild relatives may have played an adaptive role.

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
%The following is not fully clear.  Cut or revise.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
Height differences were not detected under lowland conditions.
%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

@@

Populations of \emph{mexicana} cannot be found outside of the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
%Several questions regarding \emph{mexicana} introgression yet remain.
%Has the introgression from \emph{mexicana} spread to multiple highland populations? 
%How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
%On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al. (2013) \citep{hufford2013}.
%The signal of introgression is absent for the other three populations.
%On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
%More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry. 
%The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects. 
%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
%The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
%Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
%Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
Historically, \emph{mexicana} haplotypes appear to have played an important role in adaptation of maize to challenging high-elevation conditions.
Breeding programs for highland maize may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.
